---
title: "In Class Exercise 3"
---

# Packages used

```{r}
#| eval: false
install.packages("maptools", repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

```{r}
pacman::p_load(tidyverse, sf, maptools, raster,spatstat, tmap)
```

# Spatial Data Wrangling

### Importing the spatial data

In this section ***sf*** package will be used to import these 3 geosptial data sets into R. <https://github.com/spatstat/spatstat.explore> (<https://r-spatial.github.io/sf/reference/st_read.html)>

If you stay in lang and long all the analysis will be inaccurate.

```{r}
childcare_sf <- st_read("../../data/geospatial/ChildCareServices.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz_sf <- st_read(dsn = "../../data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
plot(mpsz_sf)
```

# Creating Coastal Outline

```{r}
sg_sf <- mpsz_sf %>%
  st_union()
```

```{r}
plot(sg_sf)
```

environment: st_combine: mini points st_union: all will be combine into one point

## Geospatial Data Wrangling

### Creating ppp objects: sf methods

converting sf dataframes to sp's spatial classes. Changes to spatial point data frames. Then converting sp to ppp format.

```{r}
childcare_ppp <- as.ppp(childcare_sf)
```

chilcare_ppp is a list: Have window and the information in the data.

```{r}
summary(childcare_ppp)
```

### Handling duplicated points

We can check the duplication in ppp objects by using the code chunk (Remember: Always remember to check for duplicates in data)

```{r}
any(duplicated(childcare_ppp))
```

To count the number of co-incidence point, we will use the *mulitplicity()* function as shown below:

```{r}
multiplicity(childcare_ppp)
```

If we want to know how many locations have more than one point event:

```{r}
sum(multiplicity(childcare_ppp) > 1)
```

### Creating owin object: sf method

```{r}
sg_owin <- as.owin(sg_sf)
```

The output object can be displayed by using the *plot()* function

```{r}
plot(sg_owin)
```

Below the code in the sf way:

```{r}
pg <- mpsz_sf %>% 
  filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_sf %>% 
  filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_sf %>% 
  filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_sf %>% 
  filter(PLN_AREA_N == "JURONG WEST")
```

Plotting target planning areas

```{r}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
```

Load the relevant libraries:

```{r}
#pacman::p_load(sf, spNetwork, tmap, classInt, viridis, tidyverse)
pacman::p_load(sp, sf, rgdal, spNetwork, tmap)
```

Read the punggol data:

```{r}
network <- st_read(dsn = "../../data/geospatial", layer = "Punggol_St")
childcare <- st_read(dsn = "../../data/geospatial", layer = "Punggol_CC")
```

```{r}
childcare <-spTransform(childcare,
                        CRS("+init=epsg:3414"))
network <- spTransform(network,
                       CRS("+init=epsg:3414"))
```

We examine the structure.

Personal Note: The minimum and maximum of x and y needed to ensure we are doing the kernal density within the right WINDOW!

```{r}
tmap_mode('view')
tm_shape(childcare) + 
  tm_dots() + 
  tm_shape(network) + 
  tm_lines()
tmap_mode('plot')
```

Personal note:

-   tm_shape(childcare) + tm_dots() + will plot first before tm_shape(network) + tm_lines() is plotted. So childcare data is added as dots first before network variable's data is plotted as lines afterwards.

-   tm_map library uses LEAFLET library underneath its library so no need to create a new leaflet map yourself

-   the different variables can be selected and deselected with the layers icons - using interactive mode!

```{r}
tmap_mode('plot')
```

## Network

```{r}
lixels <- lixelize_lines(network, 750, mindist = 375)
```

```{r}
samples = lines_center(lixels)
```

On average, people are only willing to walk 750m based on a childcare centre study in NTU.So, 750 is chosen as window length for analysis. So, mindist is chosen as midpoint of 750 –\> 750/2.

TIP: OSM dataset: We can remove the pedastrians, walkpaths, cycling paths can be removed since it is unnecessary and it will increase computation time when running your notebooks.

```{r}
densities <- nkde(network,
                   events = childcare,
                   w = rep(1, nrow(childcare)),
                  samples = samples,
                   kernel_name = "quartic",
                   bw = 300,
                   div = "bw",
                   method = "simple",
                   digits = 1,
                   tol = 1,
                   grid_shape = c(1,1),
                   max_depth = 8,
                   agg = 5, #we aggregate the events within a 5m radius (faster calculation)
                   sparse = TRUE,
                   verbose = FALSE)

```

```{r}
samples$density <- densities
lixels$density <- densities

```

```{r}
# rescaling to help the mapping
samples$density <- samples$density*1000
lixels$density <- lixels$density*1000

```

```{r}
tmap_mode('view')
tm_shape(lixels) + 
  tm_lines(col = "density") +
  tm_shape(childcare) + 
  tm_dots()
tmap_mode('plot')
```

```{r}
tmap_mode('view')
tm_shape(lixels)+
  tm_lines(col="density", auto.palette.mapping=FALSE,)+
tm_shape(childcare)+
  tm_dots()
```

```{r}
tmap_mode('plot')
```
