---
title: "In Class Exercise 3"
author: "Maaruni"
execute: 
  warning: false
date: 01/15/2024
date-modified: "`r Sys.Date()`"
---

Reading the relevant spatial data:

```{r}
#childcare_sf <- st_read("../../data/geospatial/ChildCareServices.geojson") %>%
  #st_transform(crs = 3414)
```

```{r}
#sg_sf <- st_read(dsn = "../../data/geospatial", layer="CostalOutline")
```

```{r}
#mpsz_sf <- st_read(dsn = "../../data/geospatial", 
               # layer = "MP14_SUBZONE_WEB_PL")
```

### 

## Geospatial Wrangling

### Creating ppp objects: sf method

```{r}
#childcare_ppp <- as.ppp(childcare_sf)

```

```{r}
#summary(childcare_ppp)
```

Must be a point type data

```{r}
#any(duplicate(childcare_ppp_jit))
```

Creating ***owin object: sf mehthod***

```{r}
#sg_owin <- as.owin(sg_sf)
```

Personal note: need to be sf layer cannot be sp layer when converting to owin object!

The output objecct can be displayed by using *plot()* function.

```{r}
#summary(sg_owin)
```

-   polygon boundary

-   window area – unit of measurement should be meters since the other units are meters in the summary

```{r}
#pg <- mpsz_sf %>%
  #filter(PLN_AREA_N == "PUNGGOL")
#tm <- mpsz_sf %>%
  #filter(PLN_AREA_N == "TAMPINES")
#ck <- mpsz_sf %>%
  #filter(PLN_AREA_N == "CHUA CHU KANG")
#jw <- mpsz_sf %>%
  #filter(PLN_AREA_N == "JURONG WEST")
```

-   similar to sql: "select \* from table where PLN_AREA_N == "PUNGGOL"

Plotting target planning areas

```{r}
#par(mfrow=c(2,2))
#plot(pg, MAIN="PUNGGOL")
#plot(tm, MAIN="TAMPINES")
#plot(ck, MAIN="CHUA CHU KANG")
#plot(jw, MAIN="JURONG WEST")
```

-   plots will be slightly different since it is sf layer

correct one:
