---
title: "In Class Exercise 2"
author: "Maaruni"
format: 
  html: 
    code-fold: true
    code-summary: "Show the code"
    toc: true
    toc-depth: 4
execute: 
  warning: false
  eval: false
  echo: true
editor: visual
date: 01/15/2024
date-modified: "`r Sys.Date()`"
---

In this hands-on exercise, the following R packages will be used.

-   arrow,

-   lubridate,

-   tidyverse,

-   tmap,

-   sf.

Importing the relevant packages using pacman:

```{r}
#| eval: true
pacman::p_load(arrow, lubridate, tidyverse, sf, tmap, readr)
```

### Importing Grab-Posisi dataset

Read the parquet file into a dataframe. The data in the dataframe can be viewed in the 'environment' tab.

```{r}
df <- read_parquet("../../data/GrabPosisi/part-00000.parquet")
```

Convert the numerical data to POSIXct tiblr time format with the month and year more specifically stated:

```{r}
df$pingtimestamp <- as_datetime(df$pingtimestamp)
```

'datetime' is a lubridate function. '\$' sign allows you to access the 'pingtimestamp' field of the df dataframe.

## Extracting trip origin locations

```{r}
origin_df <- df %>%
  group_by(trj_id) %>%
  arrange(pingtimestamp) %>%
  filter(row_number()==1) %>%
  mutate(weekday = wday(pingtimestamp,
                        label=TRUE,
                        abbr=TRUE),
         start_hr = factor(hour(pingtimestamp)),
         day = factor(mday(pingtimestamp)))
```

-   arrange() function is the same as pd.sort() in python pandas.

❗ row_number() is generated after group_by() function is applied and its generated in all the entries within each group of trj_id.

-   wday(), hour(), mday() defines which lubridation function to use to extract the weekday, hour, day (refer to lubridate library)

-   factor() converts it to ordinal scale

## Extracting trip end locations

Own attempt:

```{r}
dest_df <- df %>%
  group_by(trj_id) %>%
  arrange(pingtimestamp) %>%
  filter(row_number()== n()) %>%
  mutate(weekday = wday(pingtimestamp,
                        label=TRUE,
                        abbr=TRUE),
         start_hr = factor(hour(pingtimestamp)),
         day = factor(mday(pingtimestamp)))
```

Correct solution uses different method of sorting pingtimestamps in descending order and getting the top entry instead of getting the last entry from ascending order in the above answer:

```{r}
destination_df <- df %>%
  group_by(trj_id) %>%
  arrange(desc(pingtimestamp)) %>%
  filter(row_number()==1) %>%
  mutate(weekday = wday(pingtimestamp,
                        label=TRUE,
                        abbr=TRUE),
         end_hr = factor(hour(pingtimestamp)),
         day = factor(mday(pingtimestamp)))
```

## Saving transformed dataframe results

Writing the tranformed dataset as RDS (R Data Serialisation)

```{r}
write_rds(origin_df, "../../data/rds/origin_df.rds")
write_rds(destination_df, "../../data/rds/destination.rds")
write_rds(destination_df, "../../data/rds/dest.rds")
```

## Re-importing previously saved work

Re-importing previously saved data:

```{r}
#| eval: false
# origin_df <- read_rds("../../data/rds/origin_df.rds")
# destination_df <- read_rds("../../data/rds/destination_df.rds")
```
