---
title: "In-Class Exercise 9: Geographical Segmentation with Spatially Constrained Clustering Techniques"
author: "Maaruni"
execute: 
  warning: false
date: 02/23/2024
date-modified: "`r Sys.Date()`"
---

# In class Exercise 10

## **The Data**

Two data sets will be used in this model building exercise, they are:

-   URA Master Plan subzone boundary in shapefile format (i.e. *MP14_SUBZONE_WEB_PL*)

-   condo_resale_2015 in csv format (i.e. *condo_resale_2015.csv*)

## **Getting Started**

The code chunks below installs and launches these R packages into R environment.

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary, ggstatsplot)
```

## **13.4 A short note about GWmodel**

[**GWmodel**](https://www.jstatsoft.org/article/view/v063i17) package provides a collection of localised spatial statistical methods, namely: GW summary statistics, GW principal components analysis, GW discriminant analysis and various forms of GW regression; some of which are provided in basic and robust (outlier resistant) forms. Commonly, outputs or parameters of the GWmodel are mapped to provide a useful exploratory tool, which can often precede (and direct) a more traditional or sophisticated statistical analysis.

## **13.5 Geospatial Data Wrangling**

### **13.5.1 Importing geospatial data**

The geospatial data used in this hands-on exercise is called MP14_SUBZONE_WEB_PL. It is in ESRI shapefile format. The shapefile consists of URA Master Plan 2014’s planning subzone boundaries. Polygon features are used to represent these geographic boundaries. The GIS data is in svy21 projected coordinates systems.

The code chunk below is used to import *MP_SUBZONE_WEB_PL* shapefile by using `st_read()` of **sf** packages.

```{r}
mpsz = st_read(dsn = "../../data/geospatial/week10", layer = "MP14_SUBZONE_WEB_PL")
```

The report above shows that the R object used to contain the imported MP14_SUBZONE_WEB_PL shapefile is called *mpsz* and it is a simple feature object. The geometry type is *multipolygon*. it is also important to note that mpsz simple feature object does not have EPSG information.

### **13.5.2 Updating CRS information**

The code chunk below updates the newly imported *mpsz* with the correct ESPG code (i.e. 3414)

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

After transforming the projection metadata, you can varify the projection of the newly transformed *mpsz_svy21* by using `st_crs()` of **sf** package.

The code chunk below will be used to varify the newly transformed *mpsz_svy21*.

```{r}
st_crs(mpsz_svy21)
```

Notice that the EPSG: is indicated as *3414* now.

Next, you will reveal the extent of *mpsz_svy21* by using `st_bbox()` of sf package.

```{r}
st_bbox(mpsz_svy21) #view extent
```

## **13.6 Aspatial Data Wrangling**

### **13.6.1 Importing the aspatial data**

The *condo_resale_2015* is in csv file format. The codes chunk below uses `read_csv()` function of **readr** package to import *condo_resale_2015* into R as a tibble data frame called *condo_resale*.

```{r}
condo_resale = read_csv("../../data/aspatial/week10/Condo_resale_2015.csv")
```

After importing the data file into R, it is important for us to examine if the data file has been imported correctly.

The codes chunks below uses `glimpse()` to display the data structure of will do the job.

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE) #see the data in XCOORD column
```

```{r}
head(condo_resale$LATITUDE) #see the data in YCOORD column
```

Next, `summary()` of base R is used to display the summary statistics of *cond_resale* tibble data frame.

```{r}
summary(condo_resale)
```

### **13.6.2 Converting aspatial data frame into a sf object**

Currently, the *condo_resale* tibble data frame is aspatial. We will convert it to a **sf** object. The code chunk below converts condo_resale data frame into a simple feature data frame by using `st_as_sf()` of **sf** packages.

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

Notice that `st_transform()` of **sf** package is used to convert the coordinates from wgs84 (i.e. crs:4326) to svy21 (i.e. crs=3414).

Personal Notes:

-   ggcorrmat: This is another package that does the same thing.

```{r}
#| fig-width: 12
#| fig-height: 12
ggcorrmat(condo_resale[,5:23])
```
