---
title: "Hands-on Exercise 3: 1st Order Spatial Point Patterns Analysis Methods & 2nd Order Spatial Point Patterns Analysis Methods"
author: "Maaruni"
execute: 
  warning: false
date: 01/13/2024
date-modified: "`r Sys.Date()`"
---

Questions we would like to explore:

-   are the childcare centres in Singapore randomly distributed throughout the country?

-   if the answer is not, then the next logical question is where are the locations with higher concentration of childcare centres?

# Getting started

```{r}
pacman::p_load(maptools, sf, raster, spatstat, tmap)
```

Please use this command in the console if unable to install maptools on mac:

```{r}
install.packages("maptools", repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

Reference: <https://stackoverflow.com/questions/77511849/i-have-obtained-some-error-trying-to-install-maptools-in-rstudio>

Reading the relevant spatial data:

```{r}
childcare_sf <- st_read("../../data/geospatial/ChildCareServices.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
sg_sf <- st_read(dsn = "../../data/geospatial", layer="CostalOutline")
```

```{r}
mpsz_sf <- st_read(dsn = "../../data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

### **Mapping the geospatial data sets**

After checking the referencing system of each geospatial data data frame, it is also useful for us to plot a map to show their spatial patterns.

```{r}
plot(sg_sf)
```

```{r}
plot(mpsz_sf)
```

Notice that all the geospatial layers are within the same map extend. This shows that their referencing system and coordinate values are referred to similar spatial context. This is very important in any geospatial analysis.

Alternatively, we can also prepare a pin map by using the code chunk below.

```{r}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
```

```{r}
tmap_mode('plot')
```

Notice that at the interactive mode, **tmap** is using [**leaflet for R**](https://rstudio.github.io/leaflet/) API. The advantage of this interactive pin map is it allows us to navigate and zoom around the map freely. We can also query the information of each simple feature (i.e. the point) by clicking of them. Last but not least, you can also change the background of the internet map layer. Currently, three internet map layers are provided. They are: ESRI.WorldGrayCanvas, OpenStreetMap, and ESRI.WorldTopoMap. The default is ESRI.WorldGrayCanvas.

##  **Geospatial Data wrangling**

Although simple feature data frame is gaining popularity again sp’s Spatial\* classes, there are, however, many geospatial analysis packages require the input geospatial data in sp’s Spatial\* classes. In this section, you will learn how to convert simple feature data frame to sp’s Spatial\* class.

### **Converting sf data frames to sp’s Spatial\* class**

The code chunk below uses [*as_Spatial()*](https://r4gdsa.netlify.app/chap04.html) of **sf** package to convert the three geospatial data from simple feature data frame to sp’s Spatial\* class.

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

Lets view the data generated:

```{r}
childcare
```

Personal observations: It has only 2 variables. Why is that so?

```{r}
mpsz
```

```{r}
sg
```

Notice that the geospatial data have been converted into their respective sp’s Spatial\* classes now.

### **Converting the Spatial\* class into generic sp format**

**spatstat** requires the analytical data in ***ppp*** object form. There is no direct way to convert a Spatial\* classes into ***ppp*** object. We need to convert the ***Spatial*** **classes**\* into ***Spatial*** object first.

The codes chunk below converts the Spatial\* classes into generic sp objects.

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

Next, you should display the sp objects properties as shown below.

```{r}
childcare_sp
```

```{r}
sg_sp
```

###  **Converting the generic sp format into spatstat’s ppp format**

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```

Now, let us plot ***childcare_ppp*** and examine the different.

```{r}
plot(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```

Notice the warning message about duplicates. In spatial point patterns analysis an issue of significant is the presence of duplicates. The statistical methodology used for spatial point patterns processes is based largely on the assumption that process are *simple*, that is, that the points cannot be coincident.

### **Handling duplicated points**

We can check the duplication in a ***ppp*** object by using the code chunk below.

```{r}
any(duplicated(childcare_ppp))
```

To count the number of co-indicence point, we will use the *multiplicity()* function as shown in the code chunk below.

```{r}
multiplicity(childcare_ppp)
```

If we want to know how many locations have more than one point event, we can use the code chunk below.

```{r}
sum(multiplicity(childcare_ppp) > 1)
```

The output shows that there are 128 duplicated point events.

To view the locations of these duplicate point events, we will plot `childcare` data by using the code chunk below.

```{r}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

```{r}
tmap_mode('plot')
```

There are three ways to overcome this problem. The easiest way is to delete the duplicates. But, that will also mean that some useful point events will be lost.

The second solution is use *jittering*, which will add a small perturbation to the duplicate points so that they do not occupy the exact same space.

The third solution is to make each point “unique” and then attach the duplicates of the points to the patterns as **marks**, as attributes of the points. Then you would need analytical techniques that take into account these marks.

The code chunk below implements the jittering approach.

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

```{r}
any(duplicated(childcare_ppp_jit))
```

### **Creating *owin* object**

When analysing spatial point patterns, it is a good practice to confine the analysis with a geographical area like Singapore boundary. In **spatstat**, an object called ***owin*** is specially designed to represent this polygonal region.

The code chunk below is used to covert *sg* SpatialPolygon object into owin object of **spatstat**.

```{r}
sg_owin <- as(sg_sp, "owin")
```
