---
title: "lalala"
---

```{r}
pacman::p_load(sf, spNetwork, tmap, classInt, viridis, tidyverse, list)
```

What are the different layers available as part of the gis dataset? Let us analyse the file structure. Lets add only the road layer with the name 'gis_osm_roads_free_1':

```{r}
# road_network <- st_read(dsn = "../../data/takehomeassigment1/malaysia-singapore-brunei-latest-free/", layer = "gis_osm_roads_free_1")
```

```{r}
# tmap_mode('view')
```

Lets find all the columns of the raod_netowrk dataset and all the the unique values for each column:

\<todo\>

try to constraint the minimum x coordinate and y coorinate and max x coordinate adn y coordinate to get only Singapore!

```{r}
# a GPKG dataset - it lives in spData package
#world_dataset <- system.file("shapes/world.gpkg", package = "spData")

#st_layers(world_dataset)
```

```{r}
  # tm_shape(road_network) + 
  #   tm_dots()
#tm_shape(road_network) + 
#tm_lines()
```

All the different file names:

```{r}
#list<-list.files('../../data/takehomeassignment1', pattern="*", full.names=TRUE)
#list
```
